### 2018-03-30 paper/3-abstractions.tex
Fix compile errors (thanks Irakli!)

\begin{itemize}
    \item \hs{store} and \hs{result} \emph{agree on inputs}, that is, for all
    input keys \hs{k}~$\in$~$I$:
    \[
    \hs{getValue}~\hs{k}~\hs{store}~\hs{==}~\hs{getValue}~\hs{k}~\hs{result}.
    \]
    In other words, no inputs were corrupted during the
    build.
    %\footnote{\todo{SLPJ}{Add the joke that GHC used to delete inputs?}}.

    \item The \hs{result} is \emph{consistent} with the \hs{task}, i.e.
    for all non-input keys \hs{k}~$\in$~$O$, the result of recomputing the
    \hs{task} matches the value stored in the \hs{result}:
    \[
    \hs{Just}~\hs{(}\hs{getValue}~\hs{k}~\hs{result)}~\hs{==}~\hs{compute}~\hs{task}~\hs{result}~\hs{k}.
    \]
\end{itemize}
A build system is \emph{correct} if it produces a correct \hs{result} for any
given \hs{task}, \hs{key} and \hs{store}.
}
\label{def-correct}
\vspace{2mm}

>>>>

\begin{itemize}
    \item \hs{result} and \hs{store} \emph{agree on inputs}, that is, for all
    input keys \hs{k}~$\in$~$I$:
    \[
    \hs{getValue}~\hs{k}~\hs{result}~\hs{==}~\hs{getValue}~\hs{k}~\hs{store}.
    \]
    In other words, no inputs were corrupted during the
    build.
    %\footnote{\todo{SLPJ}{Add the joke that GHC used to delete inputs?}}.

    \item The \hs{result} is \emph{consistent} with the \hs{task}, i.e.
    for all non-input keys \hs{k}~$\in$~$O$, the result of recomputing the
    \hs{task} matches the value stored in the \hs{result}:
    \[
    \hs{getValue}~\hs{k}~\hs{result}~\hs{==}~\hs{@\std{fromJust}@}~\hs{(}\hs{compute}~\hs{task}~\hs{result}~\hs{k)}
    \]
\end{itemize}
A build system is \emph{correct} if it produces a correct \hs{result} for any
given \hs{task}, \hs{key} and \hs{store}.
}
\label{def-correct}
\vspace{2mm}

### 2018-03-30 paper/5-implementations.tex
Fix a typo (thanks Evan!)
Fix compile errors (thanks Irakli!)

We define \Excel, with it's reordering dependency strategy, in  Fig.~\ref{fig-excel-implementation}.
\Excel's persistently stored information is a triple: (i) the dirty bit
function \hs{k}~\hs{->}~\hs{Bool}, (ii) an approximation of key dependencies
\hs{k}~\hs{->}~\hs{DependencyApproximation}~\hs{k} that \Excel uses to handle
dynamic dependencies, and (iii) the calculation chain \hs{[@@k]} recorded in the
previous build (\S\ref{sec-background-excel}).

@@

\begin{enumerate}
\item Any system using verifying traces, e.g. \Shake, is unlikely to see significant benefit from storing more than one \hs{Trace} per key\footnote{There is a small chance of a benefit if the dependencies change but the result does not, and then the dependencies change back to what they were before.}. Therefore, such systems can store \hs{Map}~\hs{k}~\hs{(Trace}~\hs{k}~\hs{v)}, where the initial \hs{k} is the \hs{key} field of \hs{Trace}.
\item Any system using \hs{Applicative} dependencies can omit the dependency keys from the \hs{Trace} as they can be recovered from the \hs{key} field.
\item Any \hs{Applicative} build system storing constructive traces, e.g. \Bazel, can index directly from the key and results to the output result~--~i.e. \hs{Map}~\hs{(@@k,}~\hs{[Hash}~\hs{v])}~\hs{(Hash}~\hs{v)}. Importantly, assuming the traces are stored on a central server, the client can compute the key and the hashes of its dependencies, then make a single call to the server to retrieve the result hash. In this formulation we have removed the possibility for a single key/dependency state to map to multiple different hashes, e.g. on a non-deterministic build~--~something \Bazel already prohibits which is discussed more in~\S\ref{sec-non-determinism}.
\item Finally, a \hs{Monad} build system with constructive traces can be stored as \hs{Map}~\hs{k}~\hs{(Choice}~\hs{k}~\hs{v)}, assuming a definition of \hs{Choice} as:
\begin{minted}[xleftmargin=10pt]{haskell}
data Choice k v = Choice k (Map (Hash v) Choice)
                | Result (Hash v)
\end{minted}
Here the \hs{Choice} encodes a tree, asking successive questions about keys, and taking different branches based on the answers, until it reaches a final result. Implementing this structure over client-server communication requires either a chatty interface with lots of round-trips per \hs{Choice} step, or sending over a part of the tree that is not subsequently explored.
\end{enumerate}

>>>>

We define \Excel, with its reordering dependency strategy, in  Fig.~\ref{fig-excel-implementation}.
\Excel's persistently stored information is a triple: (i) the dirty bit
function \hs{k}~\hs{->}~\hs{Bool}, (ii) an approximation of key dependencies
\hs{k}~\hs{->}~\hs{DependencyApproximation}~\hs{k} that \Excel uses to handle
dynamic dependencies, and (iii) the calculation chain \hs{[@@k]} recorded in the
previous build (\S\ref{sec-background-excel}).

@@

\begin{enumerate}
\item Any system using verifying traces, e.g. \Shake, is unlikely to see significant benefit from storing more than one \hs{Trace} per key\footnote{There is a small chance of a benefit if the dependencies change but the result does not, and then the dependencies change back to what they were before.}. Therefore, such systems can store \hs{Map}~\hs{k}~\hs{(Trace}~\hs{k}~\hs{v)}, where the initial \hs{k} is the \hs{key} field of \hs{Trace}.
\item Any system using \hs{Applicative} dependencies can omit the dependency keys from the \hs{Trace} as they can be recovered from the \hs{key} field.
\item Any \hs{Applicative} build system storing constructive traces, e.g. \Bazel, can index directly from the key and results to the output result~--~i.e. \hs{Map}~\hs{(@@k,}~\hs{[Hash}~\hs{v])}~\hs{(Hash}~\hs{v)}. Importantly, assuming the traces are stored on a central server, the client can compute the key and the hashes of its dependencies, then make a single call to the server to retrieve the result hash. In this formulation we have removed the possibility for a single key/dependency state to map to multiple different hashes, e.g. on a non-deterministic build~--~something \Bazel already prohibits which is discussed more in~\S\ref{sec-non-determinism}.
\item Finally, a \hs{Monad} build system with constructive traces can be stored as \hs{Map}~\hs{k}~\hs{(Choice}~\hs{k}~\hs{v)}, assuming a definition of \hs{Choice} as:
\begin{minted}[xleftmargin=10pt]{haskell}
data Choice k v = Choice k (Map (Hash v) (Choice k v))
                | Result (Hash v)
\end{minted}
Here the \hs{Choice} encodes a tree, asking successive questions about keys, and taking different branches based on the answers, until it reaches a final result. Implementing this structure over client-server communication requires either a chatty interface with lots of round-trips per \hs{Choice} step, or sending over a part of the tree that is not subsequently explored.
\end{enumerate}

### 2018-03-30 paper/6-engineering.tex
Fix compile errors (thanks Irakli!)

\noindent
Handling such tasks is possible in our framework, but requires an adjustment of
the correctness definition (\S\ref{sec-build-correctness}): instead of requiring
that the result of recomputing the \hs{task} matches the value stored in the \hs{result}:
\[
\hs{Just}~\hs{(}\hs{getValue}~\hs{k}~\hs{result)}~\hs{==}~\hs{compute}~\hs{task}~\hs{result}~\hs{k}\]
\noindent
we now require that \hs{result} contains \emph{one possible result of recomputing}
the \hs{task}:
\[
\hs{Just}~\hs{(}\hs{getValue}~\hs{k}~\hs{result)}~\hs{`elem`}~\hs{computeND}~\hs{task}~\hs{result}~\hs{k}
\]
where
\hs{computeND}~\hs{::}~\hs{Task}~\hs{MonadPlus}~\hs{k}~\hs{v}~\hs{->}~\hs{Store}~\hs{i}~\hs{k}~\hs{v}~\hs{->}~\hs{k}~\hs{->}~\hs{Maybe}~\hs{[@@v]}
returns the list of all possible results of the \hs{task} instead of just one
value (`ND' stands for `non-deterministic').
% (a possible implementation chooses
% \hs{f}~\hs{=}~\hs{ListT}~\hs{Identity}~\hs{[@@v]}).

>>>>

\noindent
Handling such tasks is possible in our framework, but requires an adjustment of
the correctness definition (\S\ref{sec-build-correctness}): instead of requiring
that the result of recomputing the \hs{task} matches the value stored in the \hs{result}:
\[
\hs{getValue}~\hs{k}~\hs{result}~\hs{==}~\hs{@\std{fromJust}@}~\hs{(}\hs{compute}~\hs{task}~\hs{result}~\hs{k)}\]
\noindent
we now require that \hs{result} contains \emph{one possible result of recomputing}
the \hs{task}:
\[
\hs{getValue}~\hs{k}~\hs{result}~\hs{`elem`}~\hs{@\std{fromJust}@}~\hs{(}\hs{computeND}~\hs{task}~\hs{result}~\hs{k)}
\]
where
\hs{computeND}~\hs{::}~\hs{Task}~\hs{MonadPlus}~\hs{k}~\hs{v}~\hs{->}~\hs{Store}~\hs{i}~\hs{k}~\hs{v}~\hs{->}~\hs{k}~\hs{->}~\hs{Maybe}~\hs{[@@v]}
returns the list of all possible results of the \hs{task} instead of just one
value (`ND' stands for `non-deterministic').
% (a possible implementation chooses
% \hs{f}~\hs{=}~\hs{ListT}~\hs{Identity}~\hs{[@@v]}).
